---
title: "Assignment 3"
author: "Maksim Nikiforov"
date: "9/2/2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Lahman)
library(tidyverse)
```

## Part 1 - Data Manipulation

### 1. 

**Q:** Install the Lahman package if you don’t have it. Read it in and print the Pitching data set out after first converting it to a tibble.

**A:** We can convert the Pitching data frame to a tibble using the `as_tibble()` function in the tidyverse library. Both the Lahman and the tidyverse libraries are initiated in the _setup_ chunk. We store in the new tibble as the object *tibblePitching*.  

```{r , echo=TRUE, eval=TRUE}
str(Pitching)

tibblePitching <- as_tibble(Pitching)
tibblePitching
```

### 2. 

**Q.** Choose only columns ending in “ID” and the IPouts, SO, and BB columns.

**A.** We're prompted to "use chaining whenever possible". Therefore, we pipe our new tibble (tibblePitching) to the `select()` function, which allows us to select some columns and omit others. We use the `ends_with()` option to choose only columns ending in “ID”.  

```{r , echo=TRUE, eval=TRUE}
tibblePitching %>% select(ends_with("ID"), IPouts, SO, BB)
```

### 3.

**Q.** Add to your code from the previous question (with chaining) and subset the data to only include rows from 1970 to 1990.

**A.** We can pipe our result from question 2 to the `filter()` function, which allows us to subset data with compound statements to only include where the _yearID_ is between 1970 and 1990 (inclusive). 

```{r , echo=TRUE, eval=TRUE}
tibblePitching %>% select(ends_with("ID"), IPouts, SO, BB) %>%
  filter(yearID >= 1970 & yearID <= 1990)
```
### 4. 

**Q.** Continuing your result, now add two new columns that are the strike out to walk ratio (SO/BB) and a ratio that is the (number of walks minus the strikeouts) divided by the innings pitched (SO-BB)/IPouts.

**A.** We pipe the result from question 3 to the `mutate()` function, which adds new columns while preserving existing ones. We name our new columns "strikeWalkRatio" and "inningsQuotient".    

```{r , echo=TRUE, eval=TRUE}
tibblePitching %>% select(ends_with("ID"), IPouts, SO, BB) %>%
  filter(yearID >= 1970 & yearID <= 1990) %>% mutate(strikeWalkRatio = SO/BB, inningsQuotient = (SO-BB)/IPouts)
```

### 5. 

**Q.** This won’t continue the chain from above. Find the average SO and average BB for each team for each of the years 2000 through 2003. Save this result as a new object.

This requires using the summarize() function in place of the mutate() function. The syntax is exactly
the same though!

**A.** We call our new object "avgSOBB" and we store in it the output of several chained functions. We pipe our tibble (tibblePitching) to the `filter0 ()` function to subset only the rows where the years span 2000 through 2003. Next, we pipe this output to be grouped by the teamID and the yearID, in ascending order. Lastly, we pipe this output to the `summarize()` function to find the average SO (meanSO) and average BB (meanBB).   

```{r , echo=TRUE, eval=TRUE}
avgSOBB <- tibblePitching %>% filter(yearID >= 2000 & yearID <= 2003) %>% group_by(teamID, yearID) %>% summarize(meanSO = mean(SO), meanBB = mean(BB))  

avgSOBB
```

### 6. 

**Q.** Take the output from question 5 and remove the average number of walks variable. Put the data into wide form with years each being a column and the average number of strikeouts being the values. Hint: You should have 30 rows and 5 columns.

**A.** We can remove the average number of walks by omitting the _meanBB_ variable from the `select()` function. We can then widen the data using `pivot_wider()`.   

```{r , echo=TRUE, eval=TRUE}
avgSO <- avgSOBB %>% select(teamID, yearID, meanSO) %>% pivot_wider(names_from = "yearID", values_from = "meanSO")

avgSO
```

## Part 2 - R Functions

### 1. 

**Q.** Write a basic function (call it `getIQR`) that takes in a vector of numbers and returns the interquartile range (75th percentile - 25th percentile). You can use the `quantile` function. Remove the name on the returned vector (i.e. make sure the returned item doesn’t have a name like 75% or something like that). Allow your function to pass unnamed arguments to the `quantile` function.

**A.** We can pass the _probs_ option to the `quantile()` function to find the differences between the 75th percentile (`probs = 0.75`) and the 25th percentile (`probs = 0.25`). We then return only the first results (`IQRval[[1]]`) to omit the name on the returned vector.  

```{r , echo=TRUE, eval=TRUE}
getIQR <- function(vector, ...) {
  IQRval <- quantile(vector, probs = 0.75, ...) - quantile(vector, probs = 0.25, ...)
  return(IQRval[[1]])
}
```

```{r, echo=TRUE, eval=TRUE}
sampleVector <- 1:10
sampleVector

getIQR(sampleVector)
```

### 2. 

**Q.** Take the output of question 5 from part 1 and apply the `getIRQ()` function to the average strike out and average walk variables (two function calls).

```{r , echo=TRUE, eval=TRUE}
getIQR(avgSOBB$meanSO)
getIQR(avgSOBB$meanBB)
```
### 3. 

**Q.** Write another function (call it `getCoefVar`) that returns the sample coefficient of variation, defined as the sample standard deviation divided by the absolute value of the sample mean. Allow the function to pass unnamed arguments to `sd` and `mean`.

**A.** We find the quotient for the sample coefficient of variation by using the `sd` and `mean` functions. We pass unnamed variables by specifying the `...` option in both the `function()` specification and within our calls to the `sd` and `mean` functions.

```{r, echo=TRUE, eval=TRUE}
getCoefVar <- function(vector, ...) {
  coefVar <- sd(vector, ...)/mean(vector, ...)
  return(coefVar)
}
```

### 4. 

**Q.** Repeat question 2 using the `getCoefVar()` function.

```{r, echo=TRUE, eval=TRUE}
getCoefVar(avgSOBB$meanSO)
getCoefVar(avgSOBB$meanBB)
```

### 5. 

**Q.** Create a function (call it `getSummaries`) that takes in a vector and returns both the IQR and the coefficient of variation. Do not rewrite your above two functions, call them. When returning your values, give them appropriate names. (This would be called a **wrapper** function.)

**A.** To return more than one thing, we will need to return a list. Lists are one-dimensional, and each element can be of a different type. We also specify named returns within the list. 

```{r, echo=TRUE, eval=TRUE}
getSummaries <- function(vector, ...) {
  IQR <- getIQR(vector, ...)
  coefVar <- getCoefVar(vector, ...)
  return(list(IQR = IQR, coefficientOfVar = coefVar))
}
```

### 6.

**Q.** Repeat question 2 with the getSummaries() function.

```{r, echo=TRUE, eval=TRUE}
getSummaries(avgSOBB$meanSO)
getSummaries(avgSOBB$meanBB)
```

### 7. 

**Q.** Why can you name a variable _sd_ in a function and not cause any issues with the `sd` function?

**A.** When we call a function in R, the function creates its own _temporary_ environment. Therefore, the local _sd_ variable and the `sd` function will live within different environments and _sd_ (the variable) will be inaccessible outside of its function.  

### 8. 

**Q.** Write your own infix function that can be used in replace of the `cbind` function (feel free to call the `cbind` function in your function :) Check that the function works on the two data frames below (df1 & df2) and then all three objects (df1, df2, vec1)!

**A.** An infix function is a function that goes between arguments (as opposed to a prefix). We can name our infix function `%combine%` and simply call `cbind()` to perform the work. 

```{r , echo=TRUE, eval=TRUE}
`%combine%` <- function(a, b) cbind(a, b)
```

```{r , echo=TRUE, eval=TRUE}
df1 <- data.frame(unif = runif(10), norm = rnorm(10))
df2 <- data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
vec1 <- rnorm(10, mean = 10)

df1 %combine% df2 %combine% vec1
cbind(df1, df2, vec1)
```